SUPPLEMENTARY INFORMATION FOR\\ NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Resolution Molecular Dynamics Simulations

### 2023-05-02 MANUSCRIPT/manuscriptSUPPL.tex
Update manuscript

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
%
\author{Peter Heftberger}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
%
\author{Matti Javanainen}
\email{matti.javanainen@helsinki.fi}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
%
\author{Jesper J. Madsen}
\affiliation{Department of Global Health, College of Public Health}
\alsoaffiliation{University of South Florida}
%
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
%
\author{Markus Miettinen}
\affiliation{MPI}
%
\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
%
\author{Georg Pabst}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
%
\author{Thomas Piggot}
\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}

@@

\begin{sidewaystable}[]
\begin{center}
    \caption{\label{SItab:params}%
    Simulation parameters used for different force fields.
    }
    \begin{tabular}{lcccc}
    \toprule
    \textbf{Parameter} & \textbf{CHARMM36} & \textbf{Slipids} & \textbf{Lipid17} & \textbf{MacRog} \\
    \midrule % NB interactions
    Neighbour list type & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} \\
    Long-range electrostatics & PME & PME & PME & PME \\
    LJ cut-off & 1.2~nm & 1.4~nm & 0.9~nm & 1.0~nm \\
    LJ modifier & Force switch 1.0--1.2~nm & -- & -- & -- \\
    Dispersion correction & -- & Energy \& pressure \cite{shirts2007accurate} & Energy \& pressure \cite{shirts2007accurate} & Energy \& pressure \cite{shirts2007accurate} \\
    \midrule % P/T control
    Thermostat & Nos\'{e}--Hoover \cite{hoover85,nose84} & Stochastic rescaling \cite{bussi07} & Nos\'{e}--Hoover \cite{hoover85,nose84} & Stochastic rescaling \cite{bussi07} \\
    Time constant ($T$) & 1~ps & 0.5~ps & 1~ps & 0.1~ps \\
    Coupling groups & Lipids \& water & Lipids \& water & Lipids \& water & Lipids \& water \\
    Barostat & Parrinello--Rahman \cite{parrinello1981polymorphic} & Berendsen \cite{berendsen1984molecular} & Parrinello--Rahman \cite{parrinello1981polymorphic} & Parrinello--Rahman \cite{parrinello1981polymorphic} \\
    Coupling type ($P$) & semi-isotropic & semi-isotropic & semi-isotropic & semi-isotropic \\
    Time constant ($P$) & 5~ps & 10~ps & 5~ps & 4~ps\\
    Compressibility & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar \\
    \midrule % Constraints
    Constraints & Bonds with H & All bonds & Bonds with H & All bonds \\
    \midrule 
    Water model & TIPS3P \cite{durell1994solvent}  & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} \\
    FF Source & CHARMM-GUI & CHARMM-GUI & CHARMM-GUI & Refs.~\citenum{kulig2016cis} \& \citenum{macrogpopc} \\
    Parameter source & CHARMM-GUI & Slipids website & CHARMM-GUI & Refs.~\citenum{kulig2016cis} \& \citenum{macrogpopc} \\
    \bottomrule
    \end{tabular}
\end{center}
\end{sidewaystable}

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.49\linewidth]{FIGS/scatt0.pdf}
    \includegraphics[width=0.49\linewidth]{FIGS/scatt5.pdf}\\
    \includegraphics[width=0.49\linewidth]{FIGS/scatt10.pdf}
    \includegraphics[width=0.49\linewidth]{FIGS/scatt15.pdf}\\
    \includegraphics[width=0.49\linewidth]{FIGS/scatt20.pdf}
    \includegraphics[width=0.49\linewidth]{FIGS/scatt25.pdf}
    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the next figure.}
    \label{SIfig:scattering1}
\end{figure}

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.49\linewidth]{FIGS/scatt30.pdf}
    \includegraphics[width=0.49\linewidth]{FIGS/scatt35.pdf}\\
    \includegraphics[width=0.49\linewidth]{FIGS/scatt40.pdf}
    \includegraphics[width=0.49\linewidth]{FIGS/scatt45.pdf}\\
    \includegraphics[width=0.49\linewidth]{FIGS/scatt50.pdf}
    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the previous figure.}
    \label{SIfig:scattering2}
\end{figure}

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=8.7cm]{../FIGS/ffminima.pdf}
    \caption{\label{SIfig:ffminima}%
    \textbf{Effecf of CHOL on the location of the first two minima in the form factor.}
    The minima are extracted from the form factors interpolated to all CHOL concentrations (Fig.~\ref{SIfig:scattering}) from experiment and simulation with the \texttt{findpeaks} function in Matlab.
    }
\end{figure}

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{../FIGS/d_vs_size.pdf}
    \caption{\label{SIfig:dvssize}%
    \textbf{Dependence of lateral diffusion coefficients on simulation box size.} The values calculated for the lipid centres of mass with \texttt{gmx msd} after eliminating leaflet drift. The values for the three system sizes are shown as markers together with fits of Eq.~\eqref{eq:pbc}. The values extrapolated to infinite system sizes are also shown in the separate column. 
    }
\end{figure}

@@

\clearpage
\subsection{Cholesterol Tilt}

\begin{figure}[htb!]
    \centering
    \includegraphics[width=8.7cm]{../FIGS/choltilt.pdf}
    \caption{\label{SIfig:choltilt}%
     \textbf{Effect of cholesterol concentration on the tilt of the cholesterol molecules.}
     The tilt is defined as the angle between the vector connecting the C3 and C17 carbons at the two ends of the rigid ring structure of cholesterol. The hydroxyl group is bound to C3 and the hydrocarbon chain to C17. The tilt distributions were gathered from the two leaflets, fitted with a Gamma distribution, and the mean values and standard deviations were extracted and shown here as the value and error estimate, respectively. The cholesterol concentration was identical between the simulations, but the points are horizontally slightly shifted for clarity.
    }
\end{figure}

@@

\section{CHARMM36 results from different simulation packages} \label{CHARMMprograms}
The results from CHARMM36 model for lipid bilayers from different 
simulation packages have been reported to give different results in
the literature \cite{piggot12,lee16}. The results are mainly
dependent on different Lennard-Jones cut-off settings, but
all the details are not quite understood. In this work we use
the results from Gromacs 5 with settings suggested to be optimal
by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
these settings to the results simulated with NAMD, OpenMM and literature
values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
that Gromacs 5 with settings suggested in webpage gives consistent
results with the literature and other simulation packages. Thus these
results are used in the main body of the paper. However, order parameters
are slightly overestimated respected to the experiments also with these
settings. \\
\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}

  \caption{\label{programsCOMP}
    Results for CHARMM36 model \cite{klauda10} from different simulation packages.
    Discussion going on at {\it https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}.
  }
\end{figure}

\clearpage
\section{Order parameter slopes as a function of cholesterol}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesEXPERIMENT.pdf}
  \caption{\label{slopes}
    Slopes of order parameters as a function of cholesterol from experiments \cite{ferreira13}.
  }
\end{figure*}

 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesBERGER.pdf}
  \caption{\label{slopesberger}
    Slopes of order parameters as a function of cholesterol from Berger simulations.
  }
\end{figure*}

 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesCHARMM.pdf}
  \caption{\label{slopescharmm}
    Slopes of order parameters as a function of cholesterol from CHARMM36 simulations.
  }
\end{figure*}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesMACROG.pdf}
  \caption{\label{slopesmacrog}
    Slopes of order parameters as a function of cholesterol from MacRog simulations.
  }
\end{figure*}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDS.pdf}
  \caption{\label{slopesslipids}
    Slopes of order parameters as a function of cholesterol from Slipids simulations.
  }
\end{figure*}
 \begin{figure*}[]
  \centering
  \includegraphics[width=19cm]{../FIGS/slopesSLIPIDST310K.pdf}
  \caption{\label{slopesslipids310K}
    Slopes of order parameters as a function of cholesterol from Slipids simulations at 298 K.
  }
\end{figure*}

\clearpage
 \section{Effect of undulations on order parameters} \label{undulations}
 \todo{This should be written based on contributions in this issue:
   https://github.com/NMRLipids/NmrLipidsCholXray/issues/16
 }

% \begin{figure}[]
%  \centering
%  \includegraphics[width=8cm]{../FIGS/slopesBERGER.pdf}
%  \caption{\label{programsCOMP}
%    Slopes of order parameters as a function of cholesterol from Berger simulations.
%  }
%\end{figure}

>>>>

% \author{Fernando Favela-Rosales}
% \affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
%
\author{Peter Heftberger}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
%
\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{matti.javanainen@helsinki.fi}
%
% \author{Jesper J. Madsen}
% \affiliation{Department of Global Health, College of Public Health}
% \alsoaffiliation{University of South Florida}
%
% \author{Josef Melcr}
% \affiliation{Institute of Organic Chemistry and Biochemistry,
% Academy of Sciences of the Czech Republic, 
% Prague 6, Czech Republic}
%
\author{Markus Miettinen}
\affiliation{Department of Chemistry, University of Bergen, Norway}
\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}

%
\author{O. H. Samuli Ollila}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{samuli.ollila@helsinki.fi}
%
\author{Georg Pabst}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
%
% \author{Thomas Piggot}
% \affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
% \alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}

@@

\begin{sidewaystable}[]
\begin{center}
    \caption{\label{SItab:params}%
    Simulation parameters used for different force fields.
    }
    \begin{tabular}{lcccc}
    \toprule
    \textbf{Parameter} & \textbf{CHARMM36} & \textbf{Slipids} & \textbf{Lipid17} & \textbf{MacRog} \\
    \midrule % NB interactions
    Neighbour list type & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} & Verlet \cite{pall2013flexible} \\
    Long-range electrostatics & PME \cite{essman95,darden93} & PME \cite{essman95,darden93} & PME \cite{essman95,darden93} & PME \cite{essman95,darden93} \\
    LJ cut-off & 1.2~nm & 1.4~nm & 0.9~nm & 1.0~nm \\
    LJ modifier & Force switch 1.0--1.2~nm & -- & -- & -- \\
    Dispersion correction & -- & Energy \& pressure \cite{shirts2007accurate} & Energy \& pressure \cite{shirts2007accurate} & Energy \& pressure \cite{shirts2007accurate} \\
    \midrule % P/T control
    Thermostat & Nos\'{e}--Hoover \cite{hoover85,nose84} & Stochastic rescaling \cite{bussi07} & Nos\'{e}--Hoover \cite{hoover85,nose84} & Stochastic rescaling \cite{bussi07} \\
    Time constant ($T$) & 1~ps & 0.5~ps & 1~ps & 0.1~ps \\
    Coupling groups & Lipids \& water & Lipids \& water & Lipids \& water & Lipids \& water \\
    Barostat & Parrinello--Rahman \cite{parrinello1981polymorphic} & Berendsen \cite{berendsen1984molecular} & Parrinello--Rahman \cite{parrinello1981polymorphic} & Parrinello--Rahman \cite{parrinello1981polymorphic} \\
    Coupling type ($P$) & semi-isotropic & semi-isotropic & semi-isotropic & semi-isotropic \\
    Time constant ($P$) & 5~ps & 10~ps & 5~ps & 4~ps\\
    Compressibility & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar & 4.5$\cdot$10$^{-5}$~1/bar \\
    \midrule % Constraints
    Constraints & Bonds with H & All bonds & Bonds with H & All bonds \\
    \midrule 
    Water model & TIPS3P \cite{durell1994solvent}  & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} & TIP3P \cite{jorgensen83} \\
    FF Source & CHARMM-GUI & CHARMM-GUI & CHARMM-GUI & Refs.~\citenum{Kulig16} \& \citenum{macrogpopc} \\
    Parameter source & CHARMM-GUI & Slipids website & CHARMM-GUI & Refs.~\citenum{Kulig16} \& \citenum{macrogpopc} \\
    \bottomrule
    \end{tabular}
\end{center}
\end{sidewaystable}

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt0.pdf}
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt5.pdf}\\
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt10.pdf}
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt15.pdf}\\
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt20.pdf}
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt25.pdf}
    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the next figure.}
    \label{SIfig:scattering1}
\end{figure}

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt30.pdf}
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt35.pdf}\\
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt40.pdf}
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt45.pdf}\\
    \includegraphics[width=0.45\linewidth]{../FIGS/scatt50.pdf}
    \caption{Scattering intensities from X-ray scattering experiments with various concentrations of cholesterol. More data are shown in the previous figure.}
    \label{SIfig:scattering2}
\end{figure}

@@

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.9\linewidth]{../FIGS/d_vs_size.pdf}
    \caption{\label{SIfig:dvssize}%
    \textbf{Dependence of lateral diffusion coefficients on simulation box size.} The values calculated for the lipid centres of mass with \texttt{gmx msd} after eliminating leaflet drift. The values for the three system sizes are shown as markers together with fits of Eq.~\eqref{eq:sdlatpbc}. The values extrapolated to infinite system sizes are also shown in the separate column. 
    }
\end{figure}

@@

@@
